---
title: "P8105 Homework 2"
author: "Lauren Holley"
output: github_document
---
```{r message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

```{r}
subways_df = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
   pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_transform = as.character
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE),
    vending = case_match(
      vending,
      "YES" ~ TRUE,
      "NO" ~ FALSE),
      route_number = as.numeric(route_number)
  ) |>
  select(line:entry, vending, ada, route_number, route_name)
```

My steps for cleaning:

1. `clean_names` to make all column names snake case.

2. `pivot_longer` since the route columns were unnecessarily separated. I added a prefix "route" so only the number of the route was retained. Those numbers were then places into `route_number`. Values of each of these columns were sent to `route_name`, and then transformed to characters due to their mixed numeric and character nature. 

3. I did the suggested `case_match` step from instructions to create a logical assignment in the `entry` variable. I then realized that `vending` was also a chr variable and switched it to logical instead.

4. I made route numbers numeric using `as.numeric`.

5. I selected for the requested variables. 

The `subways_df` dataset contains variable names: `r colnames(subways_df)`. The dataset includes `r ncol(subways_df)` columns and `r nrow(subways_df)` rows. 

```{r}
nrow(distinct(subways_df, line, station_name))
```

There are 465 distinct stations.

```{r}
subways_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

There are 84 stations that are ADA compliant.

```{r}
subways_df |>
  filter(vending == FALSE) |>
  summarize(proportion = mean(entry))
```

37.7% of station entrances / exits without vending allow entrance.

```{r}
subways_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

60 distinct stations serve the A train. 

```{r}
subways_df |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

17 stations that serve the A train are also ADA compliant.

## Problem 2

Importing the Trash Wheel data!!

```{r}
mr_trash_df =
  readxl::read_excel(
    "data/Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N709"
    ) |>
  janitor::clean_names() |>
  mutate(
    sports_balls = as.integer(round(sports_balls, digits = 0)),
    year = as.numeric(year),
    trash_wheel = "mr"
  )

prof_trash_df =
  readxl::read_excel(
    "data/Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M134"
    ) |>
  janitor::clean_names() |>
  mutate(
    trash_wheel = "professor"
  )

gwynns_trash_df =
  readxl::read_excel(
    "data/Trash Wheel Collection Data.xlsx",
    sheet = "Gwynns Falls Trash Wheel",
    range = "A2:L351"
    ) |>
  janitor::clean_names() |>
  mutate(
    trash_wheel = "gwynns"
  )
```

Binding rows: 

```{r}
trash_wheel_df =
  bind_rows(mr_trash_df, prof_trash_df, gwynns_trash_df) |>
  relocate(trash_wheel)
```

The `trash_wheel_df` dataset contains variable names: `r colnames(trash_wheel_df)`. The dataset includes `r ncol(trash_wheel_df)` columns and `r nrow(trash_wheel_df)` rows. Some key variables include `trash_wheel` with values mr, professor, and gwynns, which indicate which trash wheel the data originated from, and `weight_tons`, signifying the weight in tons of trash collected through a `<dbl>` value. 

* The total weight collected by Professor Trash Wheel was `r round(sum(pull(trash_wheel_df |> filter(trash_wheel == "professor") |> drop_na(weight_tons), weight_tons)), 2)` tons. 

* The total number of cigarette butts collected by the Gwynns Falls Trash Wheel was
`r as.integer(sum(pull(trash_wheel_df |> filter(trash_wheel == "gwynns", month == "June", year == 2022) |> drop_na(cigarette_butts), cigarette_butts)))` in June 2022.

## Problem 3

#### Baseline Dataset 

```{r}
baseline_df = (
  read_csv("data/MCI_baseline.csv",
           skip = 1,
           na = ".") |>
    janitor::clean_names() |>
    rename(bl_age = current_age) |>
    mutate(
      sex = case_match(
        sex,
        0 ~ "female",
        1 ~ "male"),
      apoe4 = case_match(
        apoe4,
        0 ~ "non_carrier",
        1 ~ "carrier")
    ) |>
    filter(is.na(age_at_onset) | age_at_onset > bl_age)
)
```

Important steps for the `baseline_df` import included the skip step to remove encoding instructions put in by the creators of the dataset, as well as coding the `.` values in the dataset to be `NA`. Other steps included using `clean_names` to make variables snake case, as well as the assignment requirements to properly encode character values for both the `sex` and `apoe4` variables. Lastly, I went back and renamed `current_age` to `bl_age` for easier personal understanding of the data. 

* In total, 483 participants were recruited, with 479 remaining within the inclusion criteria for no MCI at baseline. Of these remaining, `r baseline_df |> filter(!is.na(age_at_onset)) |> nrow()` developed MCI. 

* The average baseline age was `r round(mean(pull(baseline_df, bl_age)), 2)` years old. 

* `r baseline_df |> filter(sex == "female") |> summarize(proportion = mean(apoe4 == "carrier"))` is the proportion of women in the study that were APOE4 carriers. 

#### Longitudinal Marker Dataset

```{r}
amyloid_df = (
  read_csv("data/mci_amyloid.csv",
           skip = 1,
           na = c("NA", "Na")) |>
    janitor::clean_names() |>
    pivot_longer(
      baseline:time_8,
      names_to = "time",
      names_prefix = "time_",
      values_to = "amyloid_ratio"
    ) |>
  mutate(
    time = replace(time, time == "baseline", "0"),
    time = as.numeric(time)
  )
)
```

Important steps I took in the `amyloid_df` import (similar to the last one) included the skip step to remove encoding instructions put in by the creators of the dataset, as well as coding the `NA` and `Na` values in the dataset to be `NA`. Other steps included using `clean_names` to make variables snake case, and utilizing `pivot_longer` to reassign `time` and `amyloid_ratio` into columns. Through this process, I removed the `time_` prefixes for readability, and then changed the baseline to 0 to be able to make time a `<dbl>` type variable. The dataset includes `r ncol(amyloid_df)` columns and `r nrow(amyloid_df)` rows.