---
title: "P8105 Homework 2"
author: "Lauren Holley"
output: github_document
---
```{r message=FALSE}
library(tidyverse)
```

## Reading and cleaning the subways dataset

```{r}
subways_df = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
   pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_transform = as.character
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE),
      route_number = as.numeric(route_number)
  ) |>
  select(line:entry, vending, ada, route_number, route_name)
```

My steps for cleaning:

1. `clean_names` to make all column names snake case.

2. `pivot_longer` since the route columns were unnecessarily separated. I added a prefix "route" so only the number of the route was retained. Those numbers were then places into `route_number`. Values of each of these columns were sent to `route_name`, and then transformed to characters due to their mixed numeric and character nature. 

3. I did the suggested `case_match` step from instructions to create a logical assignment in the entry variable.

4. I made route numbers numeric using `as.numeric`.

5. I selected for the requested variables. 

The `subways_df` dataset contains variable names: `r colnames(subways_df)`. The dataset includes `r ncol(subways_df)` columns and `r nrow(subways_df)` rows. 
